Memory-Efficient Mixed-Precision Implementations\\for Robust Explicit Model Predictive Control

### 2019-04-12 emsoft-2019/RMPC_Daisy.tex

\subsection{Robust Explicit MPC}
\def\reals{\mathbb{R}}
We consider the class of linear time-invariant (LTI) systems characterized by the difference equation
\begin{equation}
\label{eq:DSS}
x_{k+1}=Ax_k+Bu_k+Ew_k,\quad  k=0,1,2,\ldots
\end{equation}
where $x\in \reals^{n\times 1}$ is the state, $u\in \reals^{m\times 1}$ is the control input and $w\in \reals^{d\times 1}$ is the disturbance. Matrices $A\in \reals^{n\times n}$, $B\in \reals^{n\times m}$ and $E\in \reals^{n\times d}$  capture respectively the effects of current state, input and disturbance on the next state. We assume the disturbance $w$ belongs to a set $\mathcal{W}$ where $\mathcal{W}$ is a polyhedral set.
In this paper, we focus on the robust formulation of MPC which at each time step
minimizes the worst-case value of an objective function with respect to the
disturbances over the control inputs. We assume the objective function is
quadratic with respect to the states and inputs.
We assume a linear translation on the input with the form
$$u_k=Kx_k+v_k$$
and synthesize $v_k$ instead of $u_k$. This choice reduces the 
conservativeness of the optimization and enlarges the set of feasible input
trajectories. The matrix $K$ is selected such that some desired property is satisfied under suitable assumptions on the system, e.g., stability if the pair $(A,B)$ is stabilizable.

%\eva{I don't understand the above sentence. Maybe split into several sentences? Where does $u_k$ come from (the sentence right now says you assume $u_k$)?}
%The modifed state evolution of the system will be $x_{k+1}=\bar Ax_k+Bv_k+Ew_k$ with $\bar A := A +BK$.
% and $v_k=\kappa(x_k)$ denotes the output of model predictive controller.  
The constrained optimization at each time step is of the form
\begin{align}
\label{eq:RMPC_prob}
J^{\ast}(x_0)=&\min_{v_0,\cdots,v_{N-1}} \max_{w_0,\cdots,w_{N-1}} \sum_{i=0}^{N-1}(x_i^TQx_i+u_i^TRu_i) + x_N^TQ_Fx_N\nonumber\\
\text{s.t.} \quad &x_{i+1}=Ax_i+Bu_i + E w_i, \quad\forall i\in\{0,1,\cdots,N-1\}\nonumber\\
\quad &u_i=Kx_i+v_i, \quad\forall i\in\{0,1,\cdots,N-1\}\nonumber\\
&u_i\in\mathcal{U},x_i\in\mathcal{X},\quad \forall w_i\in\mathcal{W},\,\,\forall i\in\{0,1,\cdots,N\},
%&x_i\in\mathcal{X},\quad \forall w_i\in\mathcal{W},\quad\forall i\in\{0,1,\cdots,N\},
\end{align}
%\Sadegh{The role of disturbance is missing in the inequalities. We should say, we want to satisfy the inequalities for all possible values of the disturbance.}
%
where $\mathcal U$ and $\mathcal X$ are polyhedral sets denoting the feasible
sets of inputs and states. Positive definite matrices $Q\in\reals^{n\times n}$ and
$Q_F\in\reals^{n\times n}$ indicate weights on the states. $R\in\reals^{m\times m}$ is
positive semidefinite and indicates a weight on the input in the objective function. $N$ denotes the length of the prediction horizon.
%\eva{What are $Q, Q_F, R$ and where do they come from?}

\begin{theorem}[\cite{delaPea:2005}]
\label{thm:EMPC}
The optimization \eqref{eq:RMPC_prob} can be translated into a multi-parametric quadratic problem which admits a closed-form solution. Furthermore, for the case that $R>0$, the controller is $u_k = K x_k + \kappa(x_k)$ with $\kappa(\cdot)$ being a continuous piecewise affine (PWA) function over polyhedral regions:
\begin{equation}
\kappa(x_k)=
\begin{cases}
F_1x_k+G_1 & \text{if $x_k\in \mathcal{R}_1$}\\
F_2x_k+G_2 & \text{if $x_k\in \mathcal{R}_2$}\\
\vdots\\
F_Px_k+G_P & \text{if $x_k\in \mathcal{R}_P$}
\end{cases} 
\end{equation}
where $\mathcal{R}_i$ is a polyhedral region determined by a set of linear inequalities $\mathcal R_i = \{x\in\mathcal X\,|\,H_ix\leq K_i\}$. 
\end{theorem}
The essential idea behind this theorem is to utilize the closed form $x_i=A^ix_0+\sum_{l=0}^{i-1}A^{i-l-1}Bu_l$ and transform the optimization \eqref{eq:RMPC_prob} into the following quadratic program:
%\begin{align}
%	&\min_{U_N}U_N^TZU_N+x_0^TVU_N+x_0^TYx_0\nonumber\\
%	&s.t.\quad JU_N\leq s+Dx_0
%\end{align} 
\begin{align}
\label{eq:multi_param_prog}
J^{\ast}(x_0)=\min_{z,\gamma}& \left[x_0^TYx_0+\frac{1}{2}z^THz+\gamma\right]\\
\text{s.t.} \quad &G_mz+g_m\gamma\leq W_m+S_mx_0,\nonumber\\
&G_cz\leq W_c+S_cx_0\nonumber
\end{align}
where $z\in \reals^{mN}$ and $\gamma\in\reals$ are decision variables, and $Y$, $H$, $G_m$, $G_c$, $S_m$, $S_c$, $W_m$, $W_c$, and $g_m$ 
are matrices of proper sizes that can be easily obtained from the original optimization \eqref{eq:RMPC_prob}  (see \cite{delaPea:2005}).

@@

\begin{figure}[t]
\begin{lstlisting}[mathescape=true,basicstyle=\small\ttfamily,morekeywords={if, then, elseif, return}]
if ($H_{1}\statevar \leq K_{1}$) then $u_{k}=F_1x_k+G_1$
elseif ($H_{2}\statevar \leq K_{2}$) then $u_{k}=F_2x_k+G_2$
elseif ($H_{3}\statevar \leq K_{3}$) then $u_{k}=F_3x_k+G_3$
...
elseif ($ H_{P}\statevar\leq K_{P}$) then $u_{k}=F_Px_k+G_P$
return $u_k$
\end{lstlisting}
\caption{Structure of an explicit MPC controller}
\label{lst:caseof}
\end{figure}
%Once the polytope was determined, computing for $F_ix+G_i$ requires even less effort as it only involves summation and multiplication.

>>>>

\subsection{Robust Explicit MPC}
\def\reals{\mathbb{R}}
We consider the class of linear time-invariant (LTI) systems characterized by the difference equation
\begin{equation}
\label{eq:DSS}
x_{k+1}=Ax_k+Bu_k+Ew_k,\quad  k=0,1,2,\ldots
\end{equation}
where $x\in \reals^{n\times 1}$ is the state, $u\in \reals^{m\times 1}$ is the control input and $w\in \reals^{d\times 1}$ is the disturbance. Matrices $A\in \reals^{n\times n}$, $B\in \reals^{n\times m}$ and $E\in \reals^{n\times d}$  capture respectively the effects of current state, input and disturbance on the next state. We assume the disturbance $w$ belongs to a set $\mathcal{W}$ where $\mathcal{W}$ is a polyhedral set.
In this paper, we focus on the robust formulation of MPC which at each time step
minimizes the worst-case value of an objective function with respect to the
disturbances over the control inputs. We assume the objective function is
quadratic with respect to the states and inputs.
We assume a linear translation on the input with the form
$$u_k=\mu x_k+v_k$$
and synthesize $v_k$ instead of $u_k$. This choice reduces the 
conservativeness of the optimization and enlarges the set of feasible input
trajectories. The matrix $\mu$ is selected such that some desired property is satisfied under suitable assumptions on the system, e.g., stability if the pair $(A,B)$ is stabilizable.

%\eva{I don't understand the above sentence. Maybe split into several sentences? Where does $u_k$ come from (the sentence right now says you assume $u_k$)?}
%The modifed state evolution of the system will be $x_{k+1}=\bar Ax_k+Bv_k+Ew_k$ with $\bar A := A +BK$.
% and $v_k=\kappa(x_k)$ denotes the output of model predictive controller.  
The constrained optimization at each time step is of the form
\begin{align}
\label{eq:RMPC_prob}
J^{\ast}(x_0)=&\min_{v_0,\cdots,v_{N-1}} \max_{w_0,\cdots,w_{N-1}} \sum_{i=0}^{N-1}(x_i^TQx_i+u_i^TRu_i) + x_N^TQ_Fx_N\nonumber\\
\text{s.t.} \quad &x_{i+1}=Ax_i+Bu_i + E w_i, \quad\forall i\in\{0,1,\cdots,N-1\}\nonumber\\
\quad &u_i=\mu x_i+v_i, \quad\forall i\in\{0,1,\cdots,N-1\}\nonumber\\
&u_i\in\mathcal{U},x_i\in\mathcal{X},\quad \forall w_i\in\mathcal{W},\,\,\forall i\in\{0,1,\cdots,N\},
%&x_i\in\mathcal{X},\quad \forall w_i\in\mathcal{W},\quad\forall i\in\{0,1,\cdots,N\},
\end{align}
%\Sadegh{The role of disturbance is missing in the inequalities. We should say, we want to satisfy the inequalities for all possible values of the disturbance.}
%
where $\mathcal U$ and $\mathcal X$ are polyhedral sets denoting the feasible
sets of inputs and states. Positive definite matrices $Q\in\reals^{n\times n}$ and
$Q_F\in\reals^{n\times n}$ indicate weights on the states. $R\in\reals^{m\times m}$ is
positive semidefinite and indicates a weight on the input in the objective function. $N$ denotes the length of the prediction horizon.
%\eva{What are $Q, Q_F, R$ and where do they come from?}

\begin{theorem}[\cite{delaPea:2005}]
\label{thm:EMPC}
The optimization \eqref{eq:RMPC_prob} can be translated into a multi-parametric quadratic problem which admits a closed-form solution. Furthermore, for the case that $R>0$, the controller is $u_k = \mu x_k + \kappa(x_k)$ with $\kappa(\cdot)$ being a continuous piecewise affine (PWA) function over polyhedral regions:
\begin{equation}
\kappa(x_k)=
\begin{cases}
F_1x_k+G_1 & \text{if $x_k\in \mathcal{R}_1$}\\
F_2x_k+G_2 & \text{if $x_k\in \mathcal{R}_2$}\\
\vdots\\
F_Px_k+G_P & \text{if $x_k\in \mathcal{R}_P$}
\end{cases} 
\end{equation}
where $\mathcal{R}_i$ is a polyhedral region determined by a set of linear inequalities $\mathcal R_i = \{x\in\mathcal X\,|\,H_ix\leq K_i\}$. 
\end{theorem}
The essential idea behind this theorem is to utilize the closed form $x_i=A^ix_0+\sum_{l=0}^{i-1}A^{i-l-1}Bu_l$ and transform the optimization \eqref{eq:RMPC_prob} into the following quadratic program:
%\begin{align}
%	&\min_{U_N}U_N^TZU_N+x_0^TVU_N+x_0^TYx_0\nonumber\\
%	&s.t.\quad JU_N\leq s+Dx_0
%\end{align} 
\begin{align}
\label{eq:multi_param_prog}
J^{\ast}(x_0)=\min_{z,\gamma}& \left[x_0^TYx_0+\frac{1}{2}z^THz+\gamma\right]\\
\text{s.t.} \quad &G_mz+g_m\gamma\leq W_m+S_mx_0,\nonumber\\
&G_cz\leq W_c+S_cx_0\nonumber
\end{align}
where $z\in \reals^{mN}$ and $\gamma\in\reals$ are decision variables, and $Y$, $H$, $G_m$, $G_c$, $S_m$, $S_c$, $W_m$, $W_c$, and $g_m$ 
are matrices of proper sizes that can be easily obtained from the original optimization \eqref{eq:RMPC_prob}  (see \cite{delaPea:2005}).

@@

\begin{figure}[t]
\begin{lstlisting}[mathescape=true,basicstyle=\small\ttfamily,morekeywords={if, then, elseif, return}]
if ($H_{1}\statevar \leq K_{1}$) then $v_{k}=F_1x_k+G_1$
elseif ($H_{2}\statevar \leq K_{2}$) then $v_{k}=F_2x_k+G_2$
elseif ($H_{3}\statevar \leq K_{3}$) then $v_{k}=F_3x_k+G_3$
...
elseif ($ H_{P}\statevar\leq K_{P}$) then $v_{k}=F_Px_k+G_P$
return $v_k$
\end{lstlisting}
\caption{Structure of an explicit MPC controller}
\label{lst:caseof}
\end{figure}
%Once the polytope was determined, computing for $F_ix+G_i$ requires even less effort as it only involves summation and multiplication.

### 2019-04-12 emsoft-2019/experiments.tex

We evaluate our complete pipeline on two benchmarks. 
The first one is the inverted pendulum problem depicted in Section \ref{sec:example}, where we set
$m=0.344$ kg, $b=0.48$ N s/m, $L=1.703$ m and $T_s=0.1$ s. The gain $K$ is
selected such the $\bar A$ has poles at $-0.1$ and $-0.5$. Moreover, we select
$N=2$, $R=1$ and $Q=Q_F=100\mathbf{\it I}$, where $\mathbf{ \it I}$ denotes the
identity matrix of proper size.

Our second benchmark is a well-known 4D example for aircraft
controller design~\cite{Kapasouris:1998}.
The control inputs for the aircraft 4-D model are the elevator and flaperon angles,
and the attack and pitch angles are the output states that need to be regulated.
The open-loop system is unstable as it has a pole with positive real part. Both
control inputs are constrained between $\pm25$ degrees. The outputs are
only constrained during the first prediction horizon. You also specify scale
factors for outputs. Using the gain $K$, the poles of $\bar A$ are placed at
$-5$, $-3$, $-1$ and $-2$. The robust MPC problem is solved for $N=2$,
$R=\mathbf{\it I}$ and $Q=Q_F=5\mathbf{\it I}$. Note that for both of the
examples, matrices $R$, $Q$ and $Q_F$ are selected such that convergence to the
origin is given more weight compared to the control effort as long as
constraints are satisfied.

>>>>

We evaluate our complete pipeline on two benchmarks. 
The first one is the inverted pendulum problem depicted in Section \ref{sec:example}, where we set
$m=0.344$ kg, $b=0.48$ N s/m, $L=1.703$ m and $T_s=0.1$ s. The gain $\mu$ is
selected such the $\bar A$ has poles at $-0.1$ and $-0.5$. Moreover, we select
$N=2$, $R=1$ and $Q=Q_F=100\mathbf{\it I}$, where $\mathbf{ \it I}$ denotes the
identity matrix of proper size.

Our second benchmark is a well-known 4D example for aircraft
controller design~\cite{Kapasouris:1998}.
The control inputs for the aircraft 4-D model are the elevator and flaperon angles,
and the attack and pitch angles are the output states that need to be regulated.
The open-loop system is unstable as it has a pole with positive real part. Both
control inputs are constrained between $\pm25$ degrees. The outputs are
only constrained during the first prediction horizon. You also specify scale
factors for outputs. Using the gain $\mu$, the poles of $\bar A$ are placed at
$-5$, $-3$, $-1$ and $-2$. The robust MPC problem is solved for $N=2$,
$R=\mathbf{\it I}$ and $Q=Q_F=5\mathbf{\it I}$. Note that for both of the
examples, matrices $R$, $Q$ and $Q_F$ are selected such that convergence to the
origin is given more weight compared to the control effort as long as
constraints are satisfied.
